Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-20 paper/uist_2016/paper.tex
almost done w/ within-subject memo

One plausible explanation of the improvement may be inherent to the encoding of episodic memory in the first place. As suggested by the levels-of-processing theory(ref), encoding efficacy is heavily dependent on the levels to which stimuli are being actively or passively processed. Not surprisingly, superficial mental engagement would produce weaker, and less enduring cognitive representation which would demand more contextually accurate and abundant cues for memory triggery, and vice versa. Oftentimes, the cognitive representation resulted from shallow processing may not even be retrievable at all as it is simply not available for retrieval.  As previously described, both of the trips revealed in our results, those that were taken farthest away from home and those that were least frequently taken, may not be special enough to engage deep memory encoding nor often enough to reinforce incidental remembrance. This means that it would be more difficult for participants to recall these trips if the provided cues are not effective enough. Since \textit{Rewind} videos provide richer and more contextually meaningful cues than static street-side images, it is reasonable that \textit{Rewind} significantly outperforms plain images as suggested by our statistical analysis. While being asked about location and trip rememberance on one of his least frequently taken route, P7 was not able to identify the location from the static image; yet he recognized the trip from \textit{Rewind} video as ``one street (in the video) helps me (him) remember".

>>>>

One plausible explanation of the improvement may be inherent to the encoding of episodic memory in the first place. As suggested by the levels-of-processing theory(ref), encoding efficacy is heavily dependent on the levels to which stimuli are being actively or passively processed. Not surprisingly, superficial mental engagement would produce weaker, and less enduring cognitive representation which would demand more contextually accurate and abundant cues for memory triggery, and vice versa. Oftentimes, the cognitive representation resulted from shallow processing may not even be retrievable at all as it is simply not available for retrieval.  As previously described, both of the trips revealed in our results, those that were taken farthest away from home and those that were least frequently taken, may not be special enough to engage deep memory encoding nor often enough to reinforce incidental remembrance. This means that it would be more difficult for participants to recall these trips if the provided cues are not effective enough. Since \textit{Rewind} videos provide richer and more contextually meaningful cues than static street-side images, it is reasonable that \textit{Rewind} significantly outperforms plain images as suggested by our statistical analysis. While being asked about location and trip rememberance on one of his least frequently taken routes, P7 was not able to identify the location from the static image at first; yet he recognized the trip from \textit{Rewind} video as ``one street (in the video) helps me (him) remember". 
%The Effects of Divided Attention on Encoding and Retrieval Processes in Human Memory 
% http://www2.psych.ubc.ca/~pgraf/Psy583Readings/Craik%20et%20al%201996.pdf
% Craik et al. 1996

The extent to which attention is dedicated while taking a trip may be another possible explanation for the observed improvement using Rewinds. This is perhaps particularly applicable to explain why participants who were self-identified regular GPS users tend to benefit more from Rewinds with episodic memory recalling. Laboratory work done by Craik et al. 1996 (ref) has demonstrated that the encoding of memory is more effective under full-attention conditions than divided-attention conditions.  This means that regular GPS users, who are often directed step-by-step on how to navigate through their environments, may be less likely to engage their attention while taking their routes, leading to less available cognitive representation for retrieval. P3, a self-identified regular GPS user who had 30\% recalling rate using static images and 80\% trip remembrance rate using Rewinds said ``I really don't pay attention to my surroundings when I'm driving. (I am) Focused (focused) on GPS and road"  Under this circumstance, it is almost certain that Rewinds would provide a more successful access to episodic memory as it provides more spatial-temporal cues for memory retrieval, by not only adjusting image attributes to reflect timely and weather changes, but also mimicing the motions of the users when they travel down a particular path.
